---
title: '36-315: Project 1 Graphics'
author: "Raphael Serrano"
date: "11/9/2018"
output: html_document
---

# Introduction

Our dataset contains information about 41396 Skytrax reviews. Each record includes a reviewer ID; the reviewer country; the date of the review; the airline; the cabin flown; an overall rating; and several sub-ratings for specific aspects of the reviewer's experience, such as cabin staff, seat comfort, and in-flight entertainment, etc.

# Reading in data

```{r, echo = FALSE, warning = FALSE, message = FALSE}
library(tidyverse)

# theme
project1_theme <- theme_bw() +
  theme(axis.text = element_text(size = 12, color = "black"),
        rect = element_rect(color = "blue"),
        text = element_text(size = 14, color = "black"),
        panel.background = element_rect(fill = "white"))

# palette
pal <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

# read in data
airline <- read_csv("airline.csv")

no.na <- function(df, cols) {
  result <- df
  for(col in cols) {
    result <- (result[!is.na(result[, col]),])
  }
  return(result)
}
```

# Boxplot of rating by cabin, facetted by airline

```{r, echo = FALSE, warning = FALSE, message = FALSE}
# boxplot
airline.box <- no.na(airline, c("cabin_flown", "overall_rating", "airline_name"))
top.airlines <- names(sort(table(airline$airline_name), decreasing = TRUE)[1:5])
airline.box$airline_name <- ifelse(airline.box$airline_name %in% top.airlines,
                                   airline.box$airline_name, "other")
ggplot(data = airline.box, aes(x = cabin_flown, y = overall_rating)) +
  geom_boxplot(aes(fill = cabin_flown)) +
  labs(
    title = "Overall Rating by Cabin and Airline",
    x = "Cabin Flown",
    y = "Overall Rating"
  ) +
  facet_grid(~ airline_name, margin = TRUE) +
  project1_theme +
  scale_fill_manual(values = pal) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

Overall, it seems as though the overall rating changes somewhat with cabin type across all airlines. As expected, the median is slightly higher for Business Class and First Class than in Economy and Premium Economy, and the first quartile is lowest for Economy.

Some airlines have better ratings on average than others across all cabin types. For instance, British Airways and Jet Airways tend to be rated higher than Air Canada Rouge and United Airlines for all four cabin types.

Spirit Airlines appears to be somewhat of an outlier among the popular airlines. Its ratings are extremely right-skew for all cabin types except for First Class.

# Scatterplot of review length versus sum of sub-ratings colored by cabin with airline point shapes

```{r, echo = FALSE, message = FALSE, warning = FALSE}
# scatterplot
airline.scatter <- no.na(airline, c("cabin_flown", "airline_name",
                                    "content"))
airline.scatter$airline_name <- ifelse(airline.scatter$airline_name %in% top.airlines,
                                   airline.scatter$airline_name, "other")
airline.scatter$review_length <- nchar(airline.scatter$content)
airline.scatter$rating_sum <- airline.scatter$seat_comfort_rating +
  airline.scatter$cabin_staff_rating +
  airline.scatter$food_beverages_rating +
  airline.scatter$inflight_entertainment_rating +
  airline.scatter$ground_service_rating +
  airline.scatter$wifi_connectivity_rating +
  airline.scatter$value_money_rating
airline.scatter <- no.na(airline.scatter, "rating_sum")
ggplot(data = airline.scatter, aes(x = rating_sum,
                                   y = review_length,
                                   color = cabin_flown)) +
  geom_point(aes(shape = airline_name)) +
  geom_smooth(method = "lm", se = TRUE) +
  labs(
    title = "Review Length vs Sum of Sub-Ratings",
    y = "Review Length (characters)",
    x = "Sum of Sub-Ratings",
    color = "Cabin Flown",
    shape = "Airline",
    caption = paste("Sub ratings: seat comfort, cabin staff,\nfood and beverages,",
                    "inflight entertainment,\nground service, wifi connectivity, value money")
  ) +
  project1_theme +
  scale_color_manual(values = pal[-1])
```

Most reviews are under 1000 characters long. The sum of sub-ratings appears to have a bimodal distribution with a mode near 0 and another mode above 30.

For most cabin types, there appears to be a negative linear relationship between review length and sum of sub-ratings. However, this relationship seems to be a positive one for passengers flying in Premium Economy. With that said, the difference is not statistically significant.

The relationships and the marginal distributions of review length and sum of sub-ratings do not appear to depend on the airline.

# Bar chart of counts for each airline colored by continent of reviewer

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(countrycode)
top15.airlines <- names(sort(table(airline$airline_name), decreasing = TRUE)[1:15])
airline.bar <- no.na(airline, c("author_country",
                                "airline_name"))
airline.bar$continent <- factor(countrycode(sourcevar = airline.bar[["author_country"]],
                                   origin = "country.name",
                                   destination = "continent"))
# region has lots of categories, would be hard to see on a poster
airline.bar$region <- factor(countrycode(sourcevar = airline.bar[["author_country"]],
                                   origin = "country.name",
                                   destination = "region"))
airline.bar <- airline.bar[airline.bar$airline_name %in% top15.airlines, ]

ggplot(airline.bar, aes(x = airline_name, fill = continent)) +
  geom_bar() +
  labs(title = "Top 15 airline review counts by author continent",
       x = "Airline") +
  project1_theme +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_manual(values = pal)
```

This is a stacked bar chart of review counts for the top 15 reviewed airlines by reviewer continent. The Americas and Europe are the most common reviewer continent. Africa is the least common with the most occurrences in Qatar Airways. Theis is followed by Asia which has a small presence in almost all airways. Oceania is similar except for being the large majority in Quantas airways.

# Word cloud of top words in Airline Reviews

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(wordcloud)
library(tm)
#library(ggrepel)

toSpace <- content_transformer(function (x , pattern ) gsub(pattern, " ", x))

airline.cloud <- no.na(airline, c("content"))

# create corpus so we can interface with tm
docs <- Corpus(VectorSource(airline.cloud$content))

# clean the text
docs <- tm_map(docs, toSpace, "/")
docs <- tm_map(docs, toSpace, "@")
docs <- tm_map(docs, toSpace, "\\|")
docs <- tm_map(docs, content_transformer(tolower))
docs <- tm_map(docs, removeNumbers)
docs <- tm_map(docs, removeWords, stopwords("english"))
docs <- tm_map(docs, removePunctuation)
docs <- tm_map(docs, stripWhitespace)

df <- data.frame(text = sapply(docs, as.character), stringsAsFactors = FALSE)

airline.reviews.words <- strsplit(paste(df$text, collapse = " "), " ")
airline.reviews.df <- as.data.frame(table(airline.reviews.words))
airline.reviews.df <- airline.reviews.df[order(-airline.reviews.df$Freq), ]

set.seed(36315)

layout(matrix(c(1, 2), nrow=2), heights=c(1, 4))
par(mar=rep(0, 4))
plot.new()
text(x=0.5, y=0.5, "Airline review common words")
wordcloud(words = airline.reviews.df$airline.reviews.words,
          freq = airline.reviews.df$Freq,
          min.freq = 1,
          max.words = 200,
          random.order = FALSE,
          rot.per = 0.35, 
          colors = pal)
#brewer.pal(8, "Dark2")
#ggplot(airline.reviews.df[1:100, ]) +
#  aes(x = 1, y = 1, size = Freq, label = airline.reviews.words) +
#  geom_text_repel(segment.size = 0, force = 100) +
#  scale_size(range = c(2, 15), guide = FALSE) +
#  scale_y_continuous(breaks = NULL) +
#  scale_x_continuous(breaks = NULL) +
#  labs(x = '', y = '') +
#  project1_theme
```

After cleaning, it is no surprise that flight is the most common word. This is followed by service, time, food, and seats. In that tier lies the word "good". This is indicative of a positive sentiment. This could be the result of self-report bias due to a positive experience. Expected words are also common such as cabin, seat, class, airline, economy, staff, flew, and airport.